Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy

### 2025-04-22 response.tex

\documentclass{article}

\usepackage{enumerate}
\usepackage{hyperref}

\begin{document}

\section*{Review BA2412-010}
\subsection*{Overall assessment:}
This paper by Wadsworth and Niemi adapts two models originally designed to forecast outpatient
ILI data in order to apply them to influenza hospitalization data. This is motivated by a switch in the
target definition of the US CDC FluSight challenge. The paper is well-written and pleasant to read.
The overall idea of leveraging models from previous FluSight cycles for the new target is appealing.
However, several points require clarification. Given the focus of the journal, the Bayesian inference
aspect should be developed more thoroughly as it currently seems a little ad hoc. Moreover, it
currently does not become clear to which degree the proposed methods offer improved forecasts
compared to existing (and potentially much simpler) models. I provide some specific comments in
the following.

\subsection*{Specific comments:}

\begin{enumerate}[1.]
\item The entire Bayesian inference scheme seems somewhat heuristic.

\begin{enumerate}[a.]
\item The ILI model is run separately and then feeds into the hospitalization model. So there is no
flow of information from the hospitalization data to the ILI model, which there would be in a fully
Bayesian procedure. Is this done on purpose and if so what is the motivation? The proposed
procedure reminds me of the "Bayesian cut" in the version described by Plummer
\href{(https://link.springer.com/article/10.1007/s11222-014-9503-z)}{(https://link.springer.com/article/10.1007/s11222-014-9503-z)}.
This may be a useful reference.
\item The authors employ a "modular Bayesian approach" mixing maximum likelihood and
Bayesian estimation. The authors should provide more ample references to the literature concerning
this type of procedure (is "modular Bayesian" a standard term?).
\item Also, it should be argued explicitly why this approach was chosen over a standard fully
Bayesian approach in the setting at hand. Was a fully Bayesian approach also explored? Which
difficulties were encountered?
\item Notably it would be helpful to comment on how this approach affects the uncertainty
propagation / overall quantification of forecast uncertainty.

\end{enumerate}

\item Several comments on the data and general modelling choices:

\begin{enumerate}[a.]
\item Influenza-like illness (ILI) is not a single pathogen, but a mix of several pathogens causing
similar symptoms. How reasonable is it to assume SIR dynamics for such an indicator? Do we
really assume that the tipping point due to susceptible depletion can be captured reasonably here? In
my understanding this would be the only reason to include an S compartment.

\item My impression is that the SIR does not really have any mechanistic meaning here and is just a
way of parameterizing curves that look qualitatively like ILI seasons. The ASG approach can be
seen as an alternative parameterization for such curves. Could the authors comment on what
practical differences exist between the two parameterizations?

\item Is the hospitalization indicator specific to influenza? My understanding is that it is, but this
should be more clear in Section 2.2. If it is influenza-specific: to which degree should we expect ILI
to be a helpful predictor given that it also contains COVID, RSV etc.? The influenza dynamics are
likely somewhat hidden in the ILI indicator.

\item Does the hospitalization indicator refer to incident (new) hospitalization or prevalence, i.e.,
bed occupancy?

\item Is there any "mechanistic" rationale behind the ARX formulation for the hospitalizations? If we
predict prevalence / bed occupancy I can sort of see why the differences between weeks would be
explained by ILI. If we predict new hospitalizations I am not sure what motivates this ARX
formulation.

\item Currently the ILI value from the same week is used in equation (7). Would there be a point in
exploring different lags?

\item On page 14 it says: "all negative values of [...] were set to 0 to reflect realistic values of
hospitalizations". Could this be handled more naturally by appropriate distributional assumptions?

\end{enumerate}

\item And some remarks on the evaluation.

\begin{enumerate}[a.]

\item It would be interesting to have comparisons of performance beyond the variations of the two
proposed models. There should definitely be some naive baseline model (e.g., a persistence
forecast) and some out-of-the-box model applied directly to the hospitalizations (e.g. seasonal
ARIMA). Given the forecasting task was taken from the CDC challenge it would also be interesting
to see how the proposed models fare against the other participants of the challenge.

\item The measure the authors call LWIS tends to give a lot of weight to forecast tasks with low
(expected) incidence as the relative differences between predictions and observations can be quite
large then. The "regular" WIS gives most weight to weeks with high incidence. So in a sense both
are complementary and it may be helpful to also discuss regular WIS a little more.

\item Why was the CRPS used in the simulation and WIS in the data example? These two are sort of
the same (WIS approximates CRPS), so things would be more coherent if one was used throughout.

\item Some assessment of calibration, e.g. interval coverage, should be added.

\end{enumerate}

\item The authors should at least discuss the possibility of sharing certain parameters across geographic
units as done e.g. in
\href{https://www.nature.com/articles/s41467-021-23234-5}{https://www.nature.com/articles/s41467-021-23234-5}
by Osthus and Moran.

\end{enumerate}

\section*{Review BA2412-010R1AE1}

The present paper proposes a novel two component framework for projecting Health 
and Human Services (HHS) hospitalization, based on hospitalization and 
influenza-like illness (ILI) data. In a first, step ILI data are modelled which 
are then used as a predictive covariate in the modelling and projection of the 
hospitalisation data. Probabilistic forecasts are assessed in a simulation study 
using proper scoring rules.

\begin{enumerate}[1.]

\item Since this paper is submitted to Bayesian analysis, the authors should 
underline more the modelling aspects and why this is an important contribution 
for the Bayesian community. What is the impact and importance of priors. Does 
the model borrow information between different model parts, propagate 
uncertainty ... In the introduction none of this is mentioned so the question 
arises whether Bayesian Analysis is the right journal.

\item As I understand one challenge of forecasting hospitalisation data is that the 
start of the collection of these data is very recent, so that a model only 
based on these data has not much historic data to be based on. Thus, the idea 
of using ILI data is natural. The authors could emphasize more and discuss 
potential alternative models that could even be included in the simulation 
study. I am missing a comparison to models that are not the newly proposed 
ones within the paper.

\item Why is it sensible to include ILI as a linear function when modelling 
hospitalisation. Please motivate more.

\item The authors provide also state level predictions. However, I do not see 
where a spatial component is included in the model. Is this only via P or are 
all states just modelled independently. For influenza I would think that it 
might be reasonable to assume that neighbouring states show similar behaviour 
in both ILI and hosptialisation so that borrowing strength between states 
within one model seems sensible. Please discuss.

\end{enumerate}

\subsection*{Specific comments:}

\begin{enumerate}[-]

\item Section 2, page 3: please be more clear on what you mean by the 
"first seven season". Provide a time frame.

\item Figure 1: I was wondering whether the R-package geofacet could be useful to 
show also a spatial trend across US states, in addition to temporal trend, for 
certain years,  similar to the second plot provided here:

https://hafen.github.io/geofacet/

I could image that the curves, onset and general behaviour, varies across the 
states. The authors refer to this a bit in the discussion. In addition I would 
prefer to also have months names in the axis labels, to avoid confusion that 
week 1 is not the first week in January.

Can the hospitalization figures also be shown on state level at least in an appendix.

\item Section 3.5: Please be clear what the strength of a Bayesian inference scheme is.

\item Equation 6): What does the tilde mean?

\item Page 11: Please motivate the choice of an autoregressive process of order 1 
more clearly. Why not use a higher order? Would this be beneficial?

\item Page 14: Please discuss the assumption that all components of $\alpha_s$ and 
$\sigma_\epsilon^2$ are assumed to be same across all seasons. Is this realistic?

\item Figure 7: Please make clear in the caption what the columns represent in an 
intuitive manner.

\item Equation 11, 12, 13, 14: Please motivate. The authors write in Section 3.4 
that the modelling of discrepancy might require careful selection of priors. 
Please underline more how this is taken care of in this manuscript.

\item Please check for typos.

\end{enumerate}
\end{document}

### 2025-06-30 abstract.tex
fix full fit

\documentclass{article}

\usepackage{PRIMEarxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage[linktocpage=true]{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
% \usepackage{lipsum}
\usepackage{fancyhdr}       % header
\usepackage{graphicx}       % graphics
\graphicspath{{media/}}     % organize your images and other figures under media/ folder

% Packages added by Spencer
\usepackage{amsmath,amsthm,amssymb,scrextend}
\usepackage{appendix}
\AtBeginEnvironment{appendices}{\renewcommand\thesection{\Alph{section}}}
\usepackage{enumerate}
\usepackage{csquotes}
% \usepackage{subcaption}
\usepackage{amssymb}
% \usepackage{bbm}
\usepackage[bb=pazo]{mathalpha}
\usepackage{bm}
\usepackage{newtxmath} %for making indicator function
\usepackage{booktabs} %for the table in the analysis section
% \usepackage{enumerate} %for the short algorithm
\usepackage{subfigure} %for making plots with multiple images
\usepackage{tabularx} %for the spacing in tabular

\pagestyle{empty}

\newcommand\Bbbbone{%
  \ifdefined\mathbbb%
    \mathbbb{1}%
  \else%
    \boldsymbol{\mathbb{1}}%
  \fi}

% \usepackage{xr}
% \externaldocument{supplementary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \usepackage[pdftex]{graphicx}
% Standard, old-style thesis
% \usepackage{isutraditional}   
% \chaptertitle
% Old-style, thesis numbering down to subsubsection
% \alternate
% \usepackage{rotating}
% Bibliography without numbers or labels
\usepackage{natbib}
% Use the following line if you want square brackets and numbering system
%\usepackage[square, numbers]{natbib}

% this package is for getting certain math scripts like the 1 needed for an indicator function
% \usepackage{bbm}

% using this package to make a better looking list
\usepackage{enumerate}
% try this one for a list as well. may be better
% \usepackage[font=itshape]{quoting}

% package for tables

% \renewcommand{\bibfont}{\setstretch{1}\selectfont}
% \setlength{\bibsep}{13.2pt}

\usepackage{chapterbib}
%\renewcommand\bibsection{\section*{\ref name}}
%\renewcommand\bibsection{\section*}
\renewcommand{\bibsection}{\section{References}}

%\bibliographystyle{apa}  % changed from apa, abbrv, unsrtnat
%\includeonly{titletoc,chapter1}
%Optional Package to add PDF bookmarks and hypertext links
% \usepackage[pdftex,hypertexnames=false,linktocpage=true]{hyperref}
\hypersetup{colorlinks=true,linkcolor=blue,anchorcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue,bookmarksnumbered=true,pdfview=FitB}

%\usepackage{titletoc}
% \usepackage{hyperref}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Packages/commands added by Jarad
\usepackage{xcolor}
\newcommand{\jarad}[1]{{\color{red} Jarad: #1}}
\newcommand{\spencer}[1]{{\color{blue} Spencer: #1}}

%Header
\pagestyle{fancy}
\thispagestyle{empty}
\rhead{ \textit{ }}

% Update your Headers here
\fancyhead[LO]{Bayesian nonlinear flu hospitalizations forecasting}
% \fancyhead[RE]{Firstauthor and Secondauthor} % Firstauthor et al. if more than 2 - must use \documentclass[twoside]{article}

%% Title
\title{Forecasting Influenza Hospitalizations Using a Bayesian Hierarchical Nonlinear Model with Discrepancy
%%%% Cite as
%%%% Update your official citation here when published 
% \thanks{\textit{\underline{Citation}}: 
% \textbf{Authors. Title. Pages.... DOI:000000/11111.}} 
}

\author{
  Spencer Wadsworth \\
  Iowa State University \\
  Ames, Iowa\\
  \texttt{sgw96@iastate.edu} \\
  %% examples of more authors
   \And
  Jarad Niemi \\
  Iowa State University \\
  Ames, Iowa\\
  \texttt{niemi@iastate.edu} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
\maketitle
\begin{abstract}
The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread. The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods. For the first several years of the competition, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalization counts. Reliable state and national hospitalization data has only been available since 2021, but for ILI the data has been available since 2010 and has been successfully forecast for several seasons.
In this manuscript, we introduce a two component modeling framework for forecasting weekly hospitalizations utilizing both hospitalization data and ILI data. The first component is for modeling ILI data using a dynamic nonlinear Bayesian hierarchical model. The second component is for modeling hospitalizations as a function of ILI. For hospitalization forecasts, ILI is first forecasted and then hospitalizations are forecast with ILI forecasts used as a linear or quadratic predictor. In a simulation study, two ILI forecast models, including one similar to the winning model for two seasons of the CDC forecast competition from Osthus et al.
% \cite[]{osthus2019dynamic} 
% 
and a nonlinear Bayesian hierarchical model from Ulloa 
% \cite[]{ulloa2019}
are compared. Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model. Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared. We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year. We also found that other modeling decisions such as the exact nonlinear function to be used in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
\end{abstract}

% \begin{abstract}
% The annual influenza outbreak leads to significant public health and economic burdens making it desirable to have prompt and accurate probabilistic forecasts of the disease spread. The United States Centers for Disease Control and Prevention (CDC) hosts annually a national flu forecasting competition which has led to the development of a variety of flu forecast modeling methods. Beginning in 2013, the target to be forecast was weekly percentage of patients with an influenza-like illness (ILI), but in 2021 the target was changed to weekly hospitalizations. Reliable hospitalization data has only been available since 2021, but ILI data has been available since 2010 and has been successfully forecast for several seasons. In this manuscript, we introduce a two component modeling framework for forecasting hospitalizations utilizing both hospitalization and ILI data. The first component is for modeling ILI data using a nonlinear Bayesian model. The second component is for modeling hospitalizations as a function of ILI. For hospitalization forecasts, ILI is first forecast then hospitalizations are forecast with ILI forecasts used as a predictor. In a simulation study, the hospitalization forecast model is assessed and two previously successful ILI forecast models are compared. Also assessed is the usefulness of including a systematic model discrepancy term in the ILI model. Forecasts of state and national hospitalizations for the 2023-24 flu season are made, and different modeling decisions are compared. We found that including a discrepancy component in the ILI model tends to improve forecasts during certain weeks of the year. We also found that other modeling decisions such as the exact nonlinear function to be used in the ILI model or the error distribution for hospitalization models may or may not be better than other decisions, depending on the season, location, or week of the forecast.
% \end{abstract}

% \begin{keyword}[class=MSC]
% \kwd[Primary ]{00X00}
% \kwd{00X00}
% \kwd[; secondary ]{00X00}
% \end{keyword}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% `\ ' is used here because TeX ignores    %%
%% spaces after text commands.              %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Every year the seasonal influenza outbreak burdens the public health system by infecting millions, causing an influx of primary care visits and hospitalizations and leading to between 290,000 and 650,000 deaths worldwide \cite[]{whoflufact2023}. Molinari et al. \cite[]{molinari2007annual} estimated the United States' annual economic burden from medical costs, loss of income, and deaths to be over \$87 billion. Accurate forecasting of infectious diseases can inform public decision making and ease the burden of an outbreak \cite[]{turtle2021accurate, lutz2019applying}.
There is a growing consensus that disease forecasts should be probabilistic in nature \cite[]{gneiting2014probabilistic, bracher2021evaluating}, and it has been shown that reporting forecast uncertainty along with predictions may lead to better decision making \cite[]{ramos2013probabilistic, joslyn2012uncertainty, winkler1971probabilistic}.

% \jarad{Test}
% \spencer{Test}

To better inform public decision making regarding the flu epidemic, in 2013 the United States Centers for Disease Control and Prevention (CDC) organized a national flu forecasting competition, also known as FluSight \cite[]{biggerstaff2016results, mathis2024evaluation,cdcfluforecasting2024}. Originally, over a dozen teams of researchers from academic and industry backgrounds participated in FluSight by contributing their own forecast models. Besides the 2020 season --or the flu season spanning the fall of 2020 and the winter of 2021-- FluSight has been operated annually and researchers outside the CDC have been invited to participate. Initially the target data for forecasts was influenza-like illness (ILI) data. ILI is the proportion of patients who meet a healthcare provider and who display flu like symptoms, and ILI data has been available at the state and national level since the 2010 flu season \cite[]{cdcfluview2023,cdc2024fluviewportal}. 
The collaborative ILI forecasting effort has led to a number of modeling developments in flu forecasting \cite[see references therein for more examples]{mcandrew2021adaptively, osthus2021multiscale, osthus2019dynamic, ulloa2019}, and in their paper's introduction, Osthus et al.  \cite[]{osthus2019dynamic} categorized the most commonly used flu forecasting models into four classes including mechanistic models based on differential equation compartmental models, agent based models based on population simulation, machine learning/regression models including data driven machine learning and statistical models, and data assimilation models which are constructed by assimilating mechanistic models into a probabilistic framework. An additional forecast model used in FluSight involves the combination of several forecasts into a single ensemble forecast, which has been shown to perform well relative to individual models \cite[]{mcandrew2021adaptively, ray2020ensemble, yamana2016superensemble}.

The administration of FluSight saw few changes during the first seven seasons, but the onset of the COVID-19 pandemic and subsequent developments for COVID-19 forecasting led to major modifications. As a result of the COVID-19 pandemic which began during the 2019 flu season, the typical flu outbreak behavior was altered between the 2019 and 2022 seasons \cite[]{mathis2024evaluation}. The COVID-19 pandemic led to the creation of the Health and Human Services (HHS) Patient Impact and Hospital Capacity Data System \cite[]{healthdata2024covidts} which contains COVID-19 and flu hospitalization data, and the COVID-19 Forecast Hub was founded. The COVID-19 Forecast Hub was based on FluSight but with certain major adjustments including how the forecast uncertainty is represented and the addition of the weekly publication of a multi-model ensemble forecast as the official forecast of the CDC \cite[]{bracher2021evaluating, Cramer2022-hub-dataset}. Using estimated quantiles for representing forecast uncertainty and creating a multi-model ensemble are both aspects of the COVID-19 Forecast Hub which were adopted by the flu forecast competition. Additionally the target of the flu forecasts changed from being ILI data to being HHS hospitalization data, which reports the number of hospitalizations due to a laboratory confirmed flu infection  \cite[]{mathis2024evaluation,healthdata2024covidts}. This is as a result of having COVID-19 cases in the population making ILI data, already only a proxy for flu behavior, more difficult to interpret.

The contribution of this manuscript is to introduce a two component framework for modeling HHS hospitalization forecasts where hospitalization data and years of ILI data are used to inform forecast models. The first modeling component is a model of ILI data and the second is a model of hospitalization data with ILI as a predictive covariate.
Herein we use ILI models similar to those in Osthus et al. \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} for ILI forecasting.
The model of Osthus et al. \cite[]{osthus2019dynamic} is a combined data assimilation and statistical regression model which involves a compartmental model in a probabilistic framework. Their model also includes an additional component for capturing a systematic discrepancy between the deterministic part of the model and the actual data, an idea which was first introduced by Kennedy and O'Hagan \cite[]{kennedy2001bayesian}. The model in Ulloa \cite[]{ulloa2019} is a Bayesian hierarchical regression model with an underlying function intended to capture the trajectory of the seasonal ILI data. Herein, we provide a framework under which discrepancy modeling may be used along with a general function modeling ILI data, and we show the effectiveness of including discrepancy modeling during certain periods of the flu season.

In line with the newer FluSight standard of forecasting hospitalizations, we model hospitalizations as a linear function of ILI. Thus forecasts produced herein target flu hospitalizations and are a mapping of ILI forecasts to hospitalizations. This allows for ILI data from many seasons to be exploited and for ILI forecasts to assist in forecasting hospitalizations, which has fewer seasons of data than ILI. Several modeling schemes and their forecasts for the 2023 flu hospitalization season are compared, and it is shown that the modeling decisions produce good forecast results for different states or times during the flu season.

In section \ref{sec:data} we review the ILI and hospitalization data provided by the CDC and targeted by FluSight. 
In section \ref{sec:functions} the modeling framework contributed by this manuscript is given. In the same section, functions similar to those used by Osthus et al. \cite[]{osthus2019dynamic} and Ulloa \cite[]{ulloa2019} are defined.
 These functions are the susceptible-infectious-recovered (SIR) compartmental model and the asymmetric Gaussian (ASG) function  respectively. Model fitting and implementation are described at the end of the section. Section \ref{sec:simulation2} includes a simulation study where four ILI forecast models and their use in forecasting hospitalizations are compared. Commonly used proper scoring rules \cite[]{gneiting2007strictly}, which are also introduced and defined in section \ref{sec:simulation2}, are used for comparing the forecasts. Forecasting of the 2023 flu outbreak along with assessment and comparison under several selected models is performed in section \ref{sec:analysis}. 
 Finally, the manuscript is concluded in section \ref{sec:conclusion} with general observations and some discussion.

 \section{Flu outbreak data} \label{sec:data}
In this section we introduce and define ILI and hospitalization data and visually evaluate the data. ILI and hospitalization data have been the object of forecasting for FluSight with ILI being the target for the first seven seasons and hospitalizations being the target since the 2022 season. Both of these data were collected at the state, territorial, and national level and were reported at least weekly. Overall the data is reported for 53 locations including the 50 US states, the District of Columbia (DC), Puerto Rico (PR), and at the US national level. We will refer to forecast targets throughout this manuscript. A target is the specific horizon, 1, 2, 3, or 4-weeks ahead, for a specific location and week during the season.

\subsection{Influenza-like illness data}
The US Outpatient Influenza-like Illness Surveillance Network (ILINet) collects information on respiratory illness from outpatient visits to health care providers. Over 3,400 outpatient health care providers in all 50 US states, PR, DC, and the US Virgin Islands report each week the total number of outpatient visits along with the number of ILI cases. An ILI case is defined as a ``fever (temperature of $100^{\circ}$F[$37.8^{\circ}$C] or greater) and a cough and/or a sore throat." Prior to the 2021 season, the definition included ``without a known cause other than influenza" \cite[]{cdcfluview2023}. Because other illnesses such as COVID-19, RSV, and the common cold may induce similar respiratory symptoms, ILI may include patients infected with some disease other than influenza. To know whether or not a sick patient is infected with influenza requires a laboratory test.

In 2013, when FluSight began, the ILI data was the object of the forecasts. The data was released publicly at HHS region levels, and forecast teams were asked to provide forecasts of several ILI targets on the regional levels including season onset, 1-4 week ahead ILI levels, and the week of peak ILI activity \cite[]{biggerstaff2016results,mcgowan2019collaborative}. Currently, the ILI data is collected by the CDC and published on an online portal for viewing at the national, HHS region, census, and state levels \cite[]{cdc2024fluviewportal}. To obtain ILI data, we used the \texttt{R} package \texttt{cdcfluview} which provides functions for downloading the data \cite[]{rudis2021cdcfluview}. Weekly ILI data from the national, HHS region, and census levels are available from the 1997 flu season until the current season. At the state level, data is available from the 2010 flu season to the current season.

The top of figure \ref{fig:us_ili} shows the ILI data at the national level for flu seasons 2010 to 2023. For most seasons there are 52 weeks listed, but for the 2010, 2015, and 2021 seasons there are 53 weeks because there were 53 Sundays during those seasons. 
To better align with the flu behavior, week 1 is set as the first week of August and week 52 or 53 is the last week in July of the following year. For example week 1 of the 2013 season corresponds to the first week of August 2013, and week 52 of the same season corresponds to the last week of July 2014. This convention is used for the remainder of this manuscript.

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.5]{Images/us_ili_seasons.png}

    \centering
    \includegraphics[scale=.4]{Images/us_hospitalizations.png}
    \caption{Percentage of outpatient visits with an influenza-like illness (ILI) in the US for seasons 2010 to 2023 (top) and weekly flu confirmed hospitalization counts at the national level for 2022 and 2023 flu seasons (bottom). Week 1 is the first week of August of the year the flu season begins and the last week of the season is the last week of July of the following year.}
    \label{fig:us_ili}
\end{figure}

Notable from the plots in the top of figure \ref{fig:us_ili} is the regular trajectory of the ILI. With the exception of season 2020, the ILI begins low at week 1 and increases as the fall and winter progress until the ILI reaches a peak. As spring progresses to summer, the ILI decreases to low values. As Osthus et al. \cite[]{osthus2019dynamic} point out, there is nearly always either a global or local peak at week 22 which typically corresponds to the week between Christmas and New Year's day. 
Whether local or global, the ILI holiday peak is generally expected and thought to be due to widespread holiday travel, school closure, or other unique social behavior   
\cite[]{ewing2017contact, garza2013effect}.
The only seasons when there was not a peak at week 22 were season 2022 where the season peak occurred particularly early and season 2020 which was greatly influenced by the COVID-19 pandemic.

\subsection{Hospitalization Data}

Hospital admission data, used as the object of FluSight forecasting for the 2022 and 2023 seasons, is based on the CDC's National Healthcare Safety Network (NHSN) dataset entitled \textit{HealthData.gov COVID-19 Reported Patient Impact and Hospital Capacity by State Timeseries}. Several targets of respiratory illnesses including COVID-19, RSV, and influenza are reported weekly by most hospitals in the US. In February 2022 it became mandatory for all hospitals to report the number of COVID-19 and influenza hospitalizations, and since then reporting of hospitalizations has become widespread. These data were updated every Wednesday and Friday according to NHSN guidelines \cite[]{healthdata2024covidts}.

The bottom of figure \ref{fig:us_ili} shows the weekly national hospitalizations for the 2022 and 2023 flu seasons. These plots show similarities to the ILI plots in the top of the figure in that at the early weeks of the season hospitalizations are low, but they increase in the fall to a peak after which they decrease until the flu outbreak ends. For both the 2022 and 2023 seasons, the hospitalizations peaked during the same week as ILI, and in 2023 that peak occurred during the holiday week 22. Additional plots of ILI data and hospitalization
data for select states are included in the supplementary material 
\cite[]{wadsworth2024bas}. These plots show flu outbreak trajectories similar to 
those in figure \ref{fig:us_ili} with activity increasing
to a peak and then decreasing.

% \begin{figure}
%     \centering
%     \includegraphics[scale=.5]{Images/us_hospitalizations.png}
%     \caption{Weekly flu confirmed hospitalization counts at the national level for 2022 (left) and 2023 (right) flu seasons}
%     \label{fig:us_hosp}
% \end{figure}

\section{ILI and hospitalization forecast modeling} \label{sec:functions}

% \jarad{Why not start with the ILI model from Osthus et al? It makes a lot more sense to me to start with what has already been done and then expand upon what they have done. Later you have a section called ILI Model, but isn't this the ILI model?}

The typical behavior of the ILI data which starts at lower values in the late summer and fall but which increases to a peak, usually in December or January, followed by a decline motivates the use of a nonlinear function which follows a similar trajectory. Compartmental models are standard mathematical models used for modelling disease outbreaks. One important compartmental model is the susceptible-infectious-recovered (SIR) model, which is used by some to model ILI data \cite[]{osthus2019dynamic, allen2017primer}. Ulloa \cite[]{ulloa2019} chose to use the asymmetric Gaussian (ASG) function to model ILI data. The SIR and ASG models may both be appropriate to describe ILI behavior over the course of a flu season, however there may also be systematic behavior not captured by either, necessitating an additional model component to capture the discrepancy. In the first part of this section, we present an ILI model similar to the model in Osthus et al. \cite[]{osthus2019dynamic}. With some generalization, the model may incorporate any appropriate nonlinear function, though the focus here is on the SIR and ASG functions which are defined.

With the aim of forecasting hospitalizations, we also introduce a linear model of hospitalizations data with ILI as a predictive covariate. To forecast hospitalizations, ILI data is first forecasted and the forecast is then plugged in as a covariate in the hospitalization model, thus producing hospitalization forecasts. The hospitalization model is also defined in this section, and the section is concluded with descriptions of selected prior distributions, model implementation, and posterior sampling.

\subsection{ILI Model} \label{sec:ili_model}

The proposed model for ILI for any location is given in (\ref{eq:ili_model}). Here $ILI_{s,w}$ is the ILI for flu season $s$ and week $w = 1, 2, ..., W$, where $W = 52$ or $W = 53$, depending on how many Sundays there are in a given season. The ILI is a proportion, so the Beta random variable is a natural selection for modeling. Under the parameterization in of the Beta distribution used in (\ref{eq:ili_model}) the expected value is $\pi_{s,w}$ and the variance is $\pi_{s,w}(1 - \pi_{s,w})/(1 + \kappa_s)$, making $\kappa_s$ a scale parameter. The nonlinear function $f_{\theta_s}(w)$ captures the trajectory of the ILI, and $\gamma_w$ is a discrepancy term for capturing the systematic patterns which $f_{\theta_s}(w)$ does not capture. 
\begin{equation}
\begin{aligned}
    \label{eq:ili_model}
    % \begin{flalign*}
        ILI_{s,w} &\overset{ind}{\sim} \text{Beta}(\pi_{s,w}\kappa_s,\; \kappa_s(1 - \pi_{s,w})) \\
        \text{logit}(\pi_{s,w}) &= f_{\theta_s}(w) + \gamma_w
    % \end{flalign*}
\end{aligned}
\end{equation}
In Osthus et al. \cite[]{osthus2019dynamic} $f_{\theta_s}(w) = \text{logit}(I_{s,w})$ where $I_{s,w}$ is the infectious compartment of the SIR model from (\ref{eq:sir_diff}) in section \ref{sec:sir_func}. In Ulloa \cite[]{ulloa2019} $f_{\theta_s}(w) = ASG_{\theta}(w)$ from (\ref{eq:asg_function}) in section \ref{sec:asg_func}. In Ulloa \cite[]{ulloa2019} modeling is done hierarchically over seasons.

\subsection{Susceptible-Infectious-Recovered (SIR) compartmental model} \label{sec:sir_func}

The SIR compartmental model is a mathematical model used for modeling disease outbreaks and was introduced by Kermack and McKendrick \cite[]{kermack1927contribution}. Since then, compartmental models have become standard for modeling infectious diseases \cite[]{allen2008mathematical}, and many extensions have been made and studied \cite[for example]{simon2020sir, allen2017primer, van2008deterministic}. The SIR mathematical model includes three compartments and assumes that at any time $t>0$ every individual in a closed population belongs to exactly one compartment. The three compartments are susceptible ($S$), infectious ($I$), and recovered ($R$), and their interaction over the course of an outbreak is described by the differential equations in (\ref{eq:sir_diff}).

\begin{equation}
    \label{eq:sir_diff}
    \frac{dS}{dt} = -\beta SI, \quad \frac{dI}{dt} = \beta S I - \delta I, \quad \frac{dR}{dt} = \delta I
\end{equation}
Here $S$, $I$, and $R$ represent the proportion of the population in each compartment such that $S + I + R = 1$ for all $t$. The trajectory is determined by the disease transmission rate $\beta > 0$ and the recovery rate $\delta > 0$. Respectively, these may be thought of as the expected proportion of susceptible individuals who will be infected by an infectious individual and the expected rate of recovery to an immune state for a newly infected person. Whether or not a disease outbreak is classified as an epidemic is determined by the initial susceptible population $S_0$, or the susceptible population at time $0$, and the parameter $\rho = \delta/
\beta$. If $S_0/\rho > 1$, the outbreak is considered an epidemic. It is non-epidemic if $S_0/\rho \leq 1$ \cite[]{osthus2019dynamic}. Figure \ref{fig:sir_traj} shows the trajectory of the three compartments of an SIR model where $S_0$ and $\rho$ were selected to match an outbreak that is an epidemic. In the case where $S_0 \leq \rho$, the trajectory for the $I$ compartment would never be increasing. The increase to a peak and subsequent decrease in the $I$ compartment of figure \ref{fig:sir_traj} suggest it is reasonable to model ILI by this compartment. Thus in modeling the ILI data, we consider the data to be analogous to the $I$ proportion of the population.

\begin{figure}[hbt!]
    \centering
    \includegraphics[scale=.6]{Images/sir_traj.png}
    \caption{Susceptible-infectious-recovered (SIR) model separated by compartments. The three compartments are the susceptible compartment (left), infectious (center), and recovered (right). In this example, $S_0/\rho > 1$.}
    \label{fig:sir_traj}
\end{figure}

>>>>

\begin{abstract}
The annual influenza outbreak leads to significant public health and economic 
burdens making it desirable to have prompt and accurate probabilistic forecasts 
of the disease spread. The United States Centers for Disease Control and 
Prevention (CDC) hosts annually a national flu forecasting competition which has 
led to the development of a variety of flu forecast modeling methods. For the 
first several years of the competition, the target to be forecast was weekly 
percentage of patients with an influenza-like illness (ILI), but in 2021 the 
target was changed to weekly hospitalization counts. Reliable state and national 
hospitalization data has only been available since 2021, but for ILI the data 
has been available since 2010 and has been successfully forecast for several 
seasons.
In this manuscript, we introduce a two component modeling framework for 
forecasting weekly hospitalizations utilizing both hospitalization data and ILI 
data. The first component is for modeling ILI data using a dynamic nonlinear 
Bayesian hierarchical model. The second component is for modeling 
hospitalizations as a function of ILI. For hospitalization forecasts, ILI is 
first forecasted and then hospitalizations are forecast with ILI forecasts used 
as a linear or quadratic predictor. In a simulation study, two ILI forecast 
models, including one similar to the winning model for two seasons of the CDC 
forecast competition from Osthus et al.
% \cite[]{osthus2019dynamic} 
% 
and a nonlinear Bayesian hierarchical model from Ulloa 
% \cite[]{ulloa2019}
are compared. Also assessed is the usefulness of including a systematic model 
discrepancy term in the ILI model. Forecasts of state and national 
hospitalizations for the 2023-24 flu season are made, and different modeling 
decisions are compared. We found that including a discrepancy component in the 
ILI model tends to improve forecasts during certain weeks of the year. We also 
found that other modeling decisions such as the exact nonlinear function to be 
used in the ILI model or the error distribution for hospitalization models may 
or may not be better than other decisions, depending on the season, location, 
or week of the forecast.
\end{abstract}
